---
title: "Analysis of Acheta domesticus mesh experiment (Nov 23)"
author: "Edouard Bessette"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
library(formatR)
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, tidy = TRUE, tidy.opts=list(width.cutoff=60),  fig.width = 6, fig.asp = 0.8, out.width = "80%")
```

# Load the necessary libraries

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(readxl)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(viridis)
library(survminer)
library(survival)
library(lmtest)
library(lme4)
library(RVAideMemoire)
library(car)
library(emmeans)
library(multcomp)
library(Hmisc)
library(glmmTMB)
library(modelsummary)
library(extrafont)
loadfonts()
viridis_colors <- viridis_pal()(3)
setwd("C:/Users/eb826/OneDrive - University of Exeter/PhD_EdouardMicrosporidia/Thesis/Chapter4_Leidyana gryllorum/Mesh experiment")
```

# Gametocysts analysis

## Import data

```{r }
gametocysts_data <- read_excel("Mesh_Gametocysts.xlsx")
colnames(gametocysts_data)
```

## Convert factors to appropriate levels

```{r }
gametocysts_data[c("Treatment", "R", "Box_Id", "Cricket_Age", "Dpi")] <- lapply(gametocysts_data[c("Treatment", "R", "Box_Id", "Cricket_Age", "Dpi")], factor)
gametocysts_data[c("Faeces","Faeces_ind")] <- lapply(gametocysts_data[c("Faeces","Faeces_ind")], as.numeric)
```

## Reorganise levels of Treatment variable and convert variables to factors

```{r }
#gametocysts_data$Treatment <- factor(gametocysts_data$Treatment,levels = c("NonTreated","Treated"))
```

## Data summary

```{r }
str(gametocysts_data)
summary(gametocysts_data)

# Calculate summary statistics for each treatment group
summary_stats <- aggregate(Gametocysts ~ Treatment, data = gametocysts_data, 
                           FUN = function(x) c(mean = mean(x), 
                                               median = median(x), 
                                               sd = sd(x), 
                                               min = min(x), 
                                               max = max(x)))
print(summary_stats)
```

## Create boxplot for treated group

```{r}
# Subset data for only 'Treated' level of Treatment
treated_data <- subset(gametocysts_data, Treatment == "Treated")
# Apply transformation to Gametocysts variable
treated_data$sq_Gametocysts <- sqrt(treated_data$Gametocysts)

# Create boxplot for the 'Treated' level of Treatment
Gametocysts_boxplot <- ggplot(treated_data, aes(x = Treatment, y = sq_Gametocysts)) +
  geom_boxplot() +
  geom_point() +
  theme_bw(base_family = "Times New Roman") +
  theme(
    axis.text = element_text(size = 12), 
    axis.title = element_text(size = 12, vjust = 0.5),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.direction = "vertical",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5)) +
  labs(title = "Gametocysts count in Treated Treatment", y = "Square root of Gametocysts")
Gametocysts_boxplot
```

## Create gametocyst curves over time

```{r}
# Calculate mean Gametocysts per dpi
mean_gametocysts <- aggregate(Gametocysts ~ Dpi + Treatment, data = treated_data, FUN = mean)
mean_gametocysts$Dpi <- as.numeric((mean_gametocysts$Dpi))

# Plot with mean Gametocysts per dpi and curves
Gametocyst_curves <- ggplot(mean_gametocysts, aes(x = Dpi, y = Gametocysts, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, shape = 15) +
  xlim(c(1,8)) +
  theme_bw(base_family = "Times New Roman") +
  theme(
    axis.text = element_text(size = 12), 
    axis.title = element_text(size = 12, vjust = 0.5),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.direction = "vertical",
    legend.position = "right",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
  labs(title = "Mean Gametocysts per Dpi for treated groups",
       x = "Dpi",
       y = "Mean Gametocysts")
Gametocyst_curves
```

```{r}
# Calculate mean Gametocysts per cricket age
mean_gametocysts1 <- aggregate(Gametocysts ~ Cricket_Age + Treatment, data = treated_data, FUN = mean)

# Add text for infection rounds
y <- max(mean_gametocysts1$Gametocysts) * 0.95
textaes <- data.frame(y = c(y, y, y, y, y, y),
                      x = c(1, 6, 13, 21, 28, 35),
                      lab = c("Inf 1", "Inf 2", "Inf 3", "Inf 4", "Inf 5", "Inf 6"))

# Plot with mean Gametocysts per cricket age and curves
Gametocyst_curves1 <- ggplot(mean_gametocysts1, aes(x = Cricket_Age, y = Gametocysts, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, shape = 15) + 
  scale_x_discrete(breaks = seq(19, 63, by = 1)) +
  theme_bw(base_family = "Times New Roman") +
  theme(
    axis.text = element_text(size = 12), 
    axis.title = element_text(size = 12),
    axis.text.x = element_text(angle = 50, size = 8),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.direction = "vertical",
    legend.position = "right",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
  labs(title = "Mean Gametocysts per Crikcet Age for treated groups",
       x = "Cricket Age (days)",
       y = "Mean Gametocysts") +
  geom_segment(aes(x = 1, y = 0, yend = y), 
               colour = "red", linetype = "dashed", linewidth = 0.5) + 
  geom_segment(aes(x = 6, y = 0, yend = y), 
               colour = "red", linetype = "dashed", linewidth = 0.5) +
  geom_segment(aes(x = 13, y = 0, yend = y), 
               colour = "red", linetype = "dashed", linewidth = 0.5) +
  geom_segment(aes(x = 21, y = 0, yend = y), 
               colour = "red", linetype = "dashed", linewidth = 0.5) + 
  geom_segment(aes(x = 28, y = 0, yend = y), 
               colour = "red", linetype = "dashed", linewidth = 0.5) +
  geom_segment(aes(x = 35, y = 0, yend = y), 
               colour = "red", linetype = "dashed", linewidth = 0.5) + # Add infection rounds 
   geom_text(mapping = aes(y = y*1.02, x = x, label = lab), hjust = -0.1, colour = "red", data = textaes, inherit.aes = FALSE)

Gametocyst_curves1
```

## General model

### Clean the data from 0 and NAs

```{r}
# Remove rows with missing values
treated_data <- na.omit(treated_data)
# Remove rows where Gametocysts equals 0
#cleaned_data_zero <- cleaned_data[cleaned_data$Gametocysts != 0, ]
# Remove unused levels
#cleaned_data <- droplevels(cleaned_data)
treated_data <- droplevels(treated_data)
treated_data$Cricket_Age <- as.numeric(treated_data$Cricket_Age)
str(treated_data)

gametocysts_data <- na.omit(gametocysts_data)
gametocysts_data <- droplevels(gametocysts_data)
str(gametocysts_data)
```

```{r }
# Look at the variable distrib
hist(treated_data$Gametocysts)

options(na.action = "na.fail") # needed for dredge()

Full_model_dpi <- glmmTMB(Gametocysts ~ Dpi + Faeces + N_ind + (1|Box_Id), data = treated_data, nbinom2(link = "log"))
diagnose(Full_model_dpi)
```

### Check models

```{r }
fixef(Full_model_dpi)
library(MuMIn)
dredge <- dredge(Full_model_dpi)

library(performance)
check_overdispersion(Full_model_dpi)
#Overdispersion test: dispersion ratio = 0.454 (p = 0.248)
```

### Test significance of each term of the model with Anova for the Dpi model

```{r }
summary(Full_model_dpi)
Anova(Full_model_dpi, type="II")

pairwise_comp_gdpi <- emmeans(Full_model_dpi, pairwise ~ Dpi, type="response")
pairwise_comp_gdpi
cld_dpi <- cld(pairwise_comp_gdpi, adjust="fdr", Letters=letters)

# Plot the emmeans object
## Remove spaces in .group 
cld_dpi$.group=gsub(" ", "", cld_dpi$.group)

# Apply the nudge amount in geom_text
plot_cld_dpi <- ggplot(cld_dpi, aes(x = Dpi, y = response, label = .group)) +
  geom_point(shape = 15, size = 3) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax =  asymp.UCL), width = 0.2, size = 0.7) +
  #geom_text(nudge_y = c(50, 65, 110, 125, 150, 140, 950, 300, 915),
            #color='black') +
  geom_text(aes(y = asymp.UCL),  # Set y = upper.CL
            nudge_y = 50,  # Adjust this value as needed
            size = 5,
            color='black') +
  theme_bw(base_family = "Times New Roman") + 
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),  
        axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12)) +
  labs(title = "Gametocysts emmeans across Dpi",
       x = "Days post infection",
       y = "Emmeans of Gametocyst counts") +
  scale_x_discrete(limits = as.character(0:8))
plot_cld_dpi
#ggsave("plot_cld_dpi.png", plot_cld_dpi, dpi = 500, width = 6, height = 4, units = "in")

pairwise_comp_gf <- emmeans(Full_model_dpi, pairwise ~ Faeces, type="response")
pairwise_comp_gf
cld_f <- cld(pairwise_comp_gf, adjust="fdr", Letters=letters)
```

# Survival analysis

## Import data

```{r }
survival_data <- read_excel("Mesh_Survival.xlsx")
colnames(survival_data)
```
## Convert factors to appropriate levels

```{r }
survival_data[c("Treatment", "R")] <- lapply(survival_data[c("Treatment", "R")], factor)
```

## Data summary

```{r }
str(survival_data)
summary(survival_data)
```

## Check mortality (sanity check)

```{r }
# Calculate total number of crickets and number of deaths by treatment
total_Adomesticus <- survival_data %>% 
  group_by(Treatment) %>% 
  summarise(total = n())

mortality_Adomesticus <- survival_data %>% 
  group_by(Treatment, Censor) %>% 
  summarise(mortality = sum(Censor == 1)) %>%
  filter(Censor == 1) # Keep only dead crickets

# Merge total and mortality data frames
Adomesticus_check <- merge(total_Adomesticus, mortality_Adomesticus, by = "Treatment", all = TRUE)

# Calculate percentage of mortality
Adomesticus_check$percent <- (Adomesticus_check$mortality / Adomesticus_check$total) * 100
print(Adomesticus_check)

# Create a data frame with the required columns
contingency_data <- data.frame(Treatment = Adomesticus_check$Treatment,
                               Alive = Adomesticus_check$total - Adomesticus_check$mortality,
                               Dead = Adomesticus_check$mortality)

# Set Treatment as row names
row.names(contingency_data) <- contingency_data$Treatment
contingency_data$Treatment <- NULL
print(contingency_data)

# Perform chi-square test
chi_square_test <- chisq.test(contingency_data)
print(chi_square_test)
```

## Surv curves by treatments

```{r }
surv <- survfit(Surv(Death_time, Censor) ~ Treatment, data = survival_data)
surv_plot <- ggsurvplot(
  fit = surv,
  conf.int = TRUE,
  #conf.int.style = "step",
  ylim=c(0,1),
  conf.int.alpha = 0.1, 
  surv.scale = "percent",
  surv.median.line = "hv",
  break.x.by = 7,
  title = "Cricket survival per treatment (with CI)",
  xlab ="Cricket age (days)",
  legend.title = "Treatment",
  legend.labs = c("Untreated","Treated"),
  #palette = c("#c4c4ff","#ffc4c4"),
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10, face = "bold"),
      legend.direction = "horizontal",
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot

# Get p-values as letters with coordinates
pvalue_letters <- surv_pvalue(surv, get_coord = TRUE)

# Add p-values as annotations
surv_plot$plot <- surv_plot$plot +
  ggplot2::annotate("text", x = pvalue_letters$pval.x, y = pvalue_letters$pval.y, label = pvalue_letters$pval.txt, size = 4, color = "black")

# Display the plot with p-values
surv_plot$plot

ggsave("surv_plot.png", surv_plot$plot, dpi = 500, width = 5, height = 4, units = "in")
```

## LT50 determined with median values

```{r }
summary(surv)$table
```

## Determine if we have a constant risk or not

```{r }
plotsurvivors(survival_data$Death_time, survival_data$Censor)
# The instantaneous risk is considered constant if the points on the survival curve are roughly aligned on a straight line.
## Here the instant risk is constant, then we opt for a survival regression model
```

## Surv reg model

### Find the best distribution for our survival data
```{r}
# Create vector of parametric distributions to test
distList <- c(distList <- c("extreme", "logistic", "gaussian", "weibull", "exponential", "rayleigh", "lognormal", "loglogistic", "t"))

# Function fits each distribution and extracts AIC, BIC, log-likelihood values
## The distribution is estimated on the event data without any covariate (~ 1)
fit_dist <- function(dist) {
  tmp <- survreg(Surv(Death_time, Censor) ~ 1, data = survival_data, dist = dist)
  c(AIC(tmp), BIC(tmp), as.numeric(logLik(tmp)))
}

# Apply above function to each distribution in the distList
results_list <- lapply(distList, fit_dist)

# Convert the above list of results to a data frame
results_df <- data.frame(t(matrix(unlist(results_list), nrow = 3)))
colnames(results_df) <- c("aic", "bic", "logLik")
rownames(results_df) <- distList

# Find the distribution with the lowest AIC, BIC, and logLik values
bestFitAIC <- rownames(results_df)[which.min(results_df$aic)]
bestFitBIC <- rownames(results_df)[which.min(results_df$bic)]
bestFitLogLik <- rownames(results_df)[which.max(results_df$logLik)]

# Print the results data frame and best fitting distributions
results_df
cat("\nBest fitting distribution using AIC:", bestFitAIC, "\n")
cat("\nBest fitting distribution using BIC:", bestFitBIC, "\n")
cat("\nBest fitting distribution using Log-Likelihood:", bestFitLogLik, "\n")
## The best distribution here is lognormal 

# We can check graphically how the distribution is fitting to the survival data
require(flexsurv)
s <- with(survival_data,Surv(Death_time,Censor))
sLnorm  <- flexsurvreg(s ~ 1,dist="lnorm",data=survival_data)
sExp  <- flexsurvreg(s ~ 1,dist="exponential",data=survival_data) 
plot(sLnorm) +
  lines(sExp, col = "blue")
```

```{r }
model_surv <- survreg(Surv(Death_time, Censor) ~ Treatment, data = survival_data, dist="lognormal") # distribution determined earlier
summary(model_surv)
RVAideMemoire::plotresid(model_surv)
Anova(model_surv, type="II")
# According to the model ANOVA, density had a significant impact on survival
```

# Emergence analysis

## Import data

```{r }
emergence_data <- read_excel("Mesh_Emergence.xlsx")
colnames(emergence_data)
```

### Convert factors to appropriate levels

```{r }
emergence_data[c("Treatment","R","Sex", "Box_ID")] <- lapply(emergence_data[c("Treatment","R","Sex", "Box_ID")],factor)
```

### Data summary

```{r }
str(emergence_data)

# Filter the data to include only rows where Censor equals 1
filtered_data <- filter(emergence_data, Censor == 1)

# Group by Treatment, Infection, and Density, and summarize the counts of Sex
emmergence_summary_table <- filtered_data %>%
  group_by(Treatment, R) %>%
  summarise(Count_of_Sex = n()) %>%
  arrange(Treatment, R)

# Print the summarized table
print(emmergence_summary_table)
```

## Calculate proportions and perform chi-square tests

```{r echo=TRUE, message=FALSE}
emergence_table <- emergence_data %>%
  group_by(Treatment) %>%
  summarise(y1 = sum(Censor == 1), #Number of emerged crickets
            y2 = sum(Censor == 0)   #Number of non-emerged crickets
  ) %>%
  mutate(y1_prop = y1 / (y1 + y2), #Proportion of emerged crickets
         y2_prop = y2 / (y1 + y2))   #Proportion of non-emerged crickets

print(emergence_table)

# Create the contingency table
contingency_table <- as.matrix(emergence_table[, c("y1", "y2")])
rownames(contingency_table) <- paste(emergence_table$Treatment)
colnames(contingency_table) <- c("Emerged", "Non-emerged")

# Perform chi-square test
chisq_result <- chisq.test(contingency_table)
print(chisq_result)

# Perform pairwise comparisons using fdr correction
pairwise_chisq <- pairwise.prop.test(x = contingency_table, p.adjust.method = "fdr")
print(pairwise_chisq)
```
## LT50 table

```{r }
surv_emergence <- survfit(Surv(Emergence_time,Censor) ~ Sex,data = emergence_data)
summary(surv_emergence)$table
```

## Surv curve emergence

```{r }
surv_emergence <- survfit(Surv(Emergence_time,Censor) ~ Treatment,data = emergence_data)
surv_plot_emergence <- ggsurvplot(
  fit = surv_emergence,
  surv.scale = "percent",
  title = "Cricket emergence across treatment",
  ylab = "Cricket emergence",
  xlab = "Cricket age (days)",
  legend.title = "Treatment",
  legend.labs = c("Non Treated","Treated"),
  palette = c("#c4c4ff","#ffc4c4"),
  xlim = c(40,62),
  break.x.by = 4,
  fun="event",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10, face = "bold"),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot_emergence$plot
#ggsave("emergence_plot.png", p1, dpi = 500, width = 4.5, height = 4, units = "in")

# Get p-values as letters with coordinates
pvalue_letters <- surv_pvalue(surv_emergence, get_coord = TRUE)

# Add p-values as annotations
surv_plot_emergence$plot <- surv_plot_emergence$plot +
  ggplot2::annotate("text", x = pvalue_letters$pval.x, y = pvalue_letters$pval.y, label = pvalue_letters$pval.txt, size = 4, color = "black")

# Display the plot with p-values
surv_plot_emergence$plot
```

## Surv curve male

```{r }
# Subset for males
df_males <- subset(emergence_data, Sex == "Male" | Sex == "NA")


surv_male <- survfit(Surv(Emergence_time,Censor) ~ Treatment,data = df_males)
surv_plot_male <- ggsurvplot(
  fit = surv_male,
  surv.scale = "percent",
  title = "Cricket male emergence",
  ylab = "Cricket emergence",
  legend = "none",
  #legend.title = "Treatment",
  #legend.labs = c("Non Treated","Treated"),
  palette = c("#c4c4ff","#ffc4c4"),
  xlim = c(40,62),
  break.x.by = 4,
  fun="event",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot_male
#ggsave("emergence_male_plot.png", surv_plot_male$plot, dpi = 500, width = 2.6, height = 3.5, units = "in")
```

## Surv curve female

```{r }
# Subset for females
df_females <- subset(emergence_data, Sex == "Female" | Sex == "NA")

surv_female <- survfit(Surv(Emergence_time,Censor) ~ Treatment,data = df_females)
surv_plot_female <- ggsurvplot(
  fit = surv_female,
  surv.scale = "percent",
  title = "Cricket female emergence",
  ylab = "Cricket emergence",
  legend = "none",
  #legend.title = "Treatment",
  #legend.labs = c("Non Treated","Treated"),
  palette = c("#c4c4ff","#ffc4c4"),
  xlim = c(40,62),
  break.x.by = 4,
  fun="event",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot_female
#ggsave("emergence_female_plot.png", surv_plot_female$plot, dpi = 500, width = 2.6, height = 3.5, units = "in")
```

## Arrange the surv plots together

```{r }
splots <- list(surv_plot_female, surv_plot_male)
arrange <- arrange_ggsurvplots(splots, print = TRUE, ncol = 2, nrow = 1)
arrange
```

## Generalized linear model – binomial family

```{r }
# Filter out NA values from the initial data set
emergence_data_filtered <- emergence_data %>% filter(Sex != "NA")

glm_model <- glm(Emergence_time ~ Treatment * Sex, data = (emergence_data_filtered),family = Gamma(link = "inverse"))
```

### Check model and Anova

```{r }
RVAideMemoire::plotresid(glm_model)

summary(glm_model)
Anova(glm_model,type="II")
# Only Sex is significant, with Infection and Infection:Density almost significant (0.09)
```

### Pairwise comparisons 

```{r }
pairwise_comp_trt <- emmeans(glm_model, pairwise ~ Treatment, type="response")
pairwise_comp_trt
cld(pairwise_comp_trt, adjust="fdr", Letters=letters)
```

# Feed and feces analysis

## Import data

```{r }
feed_feces_data <- read_excel("Mesh_Gametocysts.xlsx")
colnames(feed_feces_data)
summary(feed_feces_data)
```

## Reorganise levels of Treatment variable and convert variables to factors

```{r }
feed_feces_data[c("Treatment", "R", "Box_Id", "Cricket_Age", "Dpi")] <- lapply(feed_feces_data[c("Treatment", "R", "Box_Id", "Cricket_Age", "Dpi")], factor)
feed_feces_data[c("Faeces","Faeces_ind")] <- lapply(feed_feces_data[c("Faeces","Faeces_ind")], as.numeric)

# Remove rows with NA in Faeces_ind or Treatment
feed_feces_data <- feed_feces_data %>% drop_na(Faeces_ind, Treatment)

str(feed_feces_data)
```

## Calculate mean and standard deviation for each response variable by treatment

```{r }
summary_data_f <- aggregate(cbind(Faeces_ind) ~ Treatment, feed_feces_data, FUN = function(x) c(mean = mean(x), sd = sd(x)))
print(summary_data_f)
```

## Boxplot of Feed Consumed by Treatment

```{r }
boxplot_feces <- ggplot(feed_feces_data, aes(x = Treatment, y = Faeces_ind, color = Treatment)) +
  geom_boxplot() +
  geom_point() +
  scale_color_viridis_d() +
  labs(title = "Faeces produced by treatment",
       x = "Treatment",
       y = "Faeces produced per individual and per day (mg)") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
boxplot_feces
```

## Feces model

```{r }
hist(feed_feces_data$Faeces_ind)

model_feces <- lmer(Faeces_ind ~  Treatment + (1|Box_Id), data = feed_feces_data) #The box id is a random effect (different box measured across different time points)
model_feces_age <- lmer(Faeces_ind ~  Treatment * Cricket_Age + (1|Box_Id), data = feed_feces_data)

anova(model_feces, model_feces_age)
# The model including Cricket age has significant lower AIC and BIC and a higher logLik
```

#### Check residuals and compare the models and test significance of each term of the model with Anova

```{r }
RVAideMemoire::plotresid(model_feces_age)

summary(model_feces_age)
Anova(model_feces_age, type="II")
```

#### Pairwise comparisons

```{r }
pairwise_comp_faeces <- emmeans(model_feces_age, pairwise ~ Cricket_Age, type="response")
pairwise_comp_faeces
cld(pairwise_comp_faeces, adjust="fdr", Letters=letters)

pairwise_comp_faeces1 <- emmeans(model_feces_age, pairwise ~ Treatment*Cricket_Age, type="response")
pairwise_comp_faeces1

# Compute the CLDs
cld_results <- cld(pairwise_comp_faeces1, adjust="fdr", Letters=letters)

# Create the interaction plot
emmip_plot <- emmip(pairwise_comp_faeces1, Treatment ~ Cricket_Age, CIs = TRUE) +
  theme_bw(base_family = "Times New Roman") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Cricket age (days)", y = "Emmeans of prouced faeces per day (mg)")  # Add this line
emmip_plot
#ggsave("emmeans_faeces_age_plot.png", emmip_plot, dpi = 500, width = 7.4, height = 4, units = "in")
```

# Haemocytes analysis

## Import data

```{r }
haemocytes_data <- read_excel("Mesh_Hemocytes.xlsx", sheet = 2)
colnames(haemocytes_data)
summary(haemocytes_data)
```

## Reorganise levels of Treatment variable and convert variables to factors

```{r }
haemocytes_data[c("Treatment", "R", "Gregarine_pres", "Ind")] <- lapply(haemocytes_data[c("Treatment", "R", "Gregarine_pres", "Ind")], factor)
haemocytes_data[c("Weight", "Haemocyte_uL", "Haemocyte")] <- lapply(haemocytes_data[c("Weight", "Haemocyte_uL", "Haemocyte")], as.numeric)
str(haemocytes_data)
```

## Calculate mean and standard deviation for each response variable by treatment

```{r }
summary_data_f <- aggregate(cbind(Haemocyte_uL) ~ Treatment, haemocytes_data, 
                            FUN = function(x) c(mean = mean(x), se = sd(x)/sqrt(length(x))))
print(summary_data_f)
```

## Correlation Analysis
```{r }
# Remove rows with NA in Weight column
haemocytes_data_clean <- haemocytes_data %>% drop_na(Weight)

# Calculate correlation
cor(haemocytes_data_clean %>% select_if(is.numeric))
```

## GLM model
```{r }
hist(haemocytes_data_clean$Haemocyte_uL)

glm_model_treatment <- glm.nb(Haemocyte_uL ~ Treatment, data = (haemocytes_data_clean))

glm_model_treatment <- glmmTMB(Haemocyte_uL ~ Treatment, data = (haemocytes_data_clean), nbinom2(link = "log"))

# Check for overdispersion
overdispersion <- glm_model_treatment$deviance / glm_model_treatment$df.residual
print(overdispersion)
```

### Check model and Anova

```{r }
RVAideMemoire::plotresid(glm_model_treatment)

summary(glm_model_treatment)
Anova(glm_model_treatment,type="II")
```

### Pairwise comparisons

```{r }
pairwise_comp_h <- emmeans(glm_model_treatment, pairwise ~ Treatment, type="response")
pairwise_comp_h
cld(pairwise_comp_h, adjust="fdr", Letters=letters)
```

## Weight analysis from hemocytes data

```{r}
boxplot_weight <- ggplot(haemocytes_data_clean, aes(x = Treatment, y = Weight, color = Treatment),
  legend.title = "Treatment") +
  geom_boxplot() +
  geom_point() +
  scale_color_viridis_d() +
  labs(title = "Weight of sampled males by treatment",
       x = "Treatment",
       y = "Weight (mg)") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
boxplot_weight
ggsave("boxplot_weight.png", boxplot_weight, dpi = 500, width = 6, height = 4, units = "in")

# Calculate summary statistics for each treatment group with standard error
summary_stats_weight <- aggregate(haemocytes_data_clean$Weight ~ Treatment, data = haemocytes_data_clean, 
                                  FUN = function(x) c(mean = mean(x), 
                                                      median = median(x), 
                                                      se = sd(x) / sqrt(length(x)), 
                                                      min = min(x), 
                                                      max = max(x)))

# Convert the resulting matrix into a data frame for better readability
summary_stats_weight <- do.call(data.frame, summary_stats_weight)

# Give meaningful names to the columns
names(summary_stats_weight) <- c("Treatment", "Mean", "Median", "SE", "Min", "Max")

# View the summary statistics
summary_stats_weight

# Model analysis
hist(haemocytes_data_clean$Weight)
lm_weigth <- lm(haemocytes_data_clean$Weight ~ Treatment, data = haemocytes_data_clean)
summary(lm_weigth)
RVAideMemoire::plotresid(lm_weigth)
Anova(lm_weigth, type="II")
```

# Create a table that compiles all the above statistical outputs from the Anova objects

* Function to extract results

```{r}
get.univariate.results <- function(model) {
  anova_results <- car::Anova(model, type = "II") %>% broom::tidy()
  
  if ("lme" %in% class(model) || "lmerMod" %in% class(model) || "glmerMod" %in% class(model)) {
    random_effects <- paste(names(VarCorr(model)), collapse = ", ")
    variances <- paste(format(VarCorr(model), digits = 2), collapse = ", ")
  } else {
    random_effects <- NA
    variances <- NA
  }
  
  cbind(
    `Response (y)` = deparse(summary(model)$call$formula[[2]]),
    `Model` = as.character(summary(model)$call[[1]]),
    Formula = paste ("y ~", deparse(summary(model)$call$formula[[3]])),
    anova_results %>%
      dplyr::mutate(
        statistic = round(statistic, digits = 2),
        signif = gtools::stars.pval(p.value),
        p.value = round(p.value, digits = 3),
        p.value = format.pval(p.value, digits = 3)
      ),
    random_effects = random_effects,
    variances = variances
  ) %>% as.data.frame
}
```

* Create a list that contains all models

```{r}
sum.model <- list(Full_model_dpi, glm_model, model_feces_age, glm_model_treatment)
```

* Apply the function to the list

```{r message=FALSE, warning=FALSE}
univariate.output <- purrr::map(.x = sum.model, .f = get.univariate.results) %>%
  dplyr::bind_rows()
univariate.output <- dplyr::rename(univariate.output, c(Factor = term, `Test value` = statistic, `P value`= p.value))

knitr::kable(univariate.output)
write.table(univariate.output, file = "models_stat_meshexp.txt", sep = ";", dec = ".")
```