---
title: "Analysis of Acheta domesticus density experiment (Sept 23)"
author: "Edouard Bessette"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
library(formatR)
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, tidy = TRUE, tidy.opts=list(width.cutoff=60),  fig.width = 6, fig.asp = 0.8, out.width = "80%")
```

# Load the necessary libraries

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(readxl)
library(dplyr)
library(ggplot2)
library(viridis)
library(survminer)
library(survival)
library(lmtest)
library(lme4)
library(RVAideMemoire)
library(car)
library(emmeans)
library(multcomp)
library(Hmisc)
library(glmmTMB)
library(extrafont)
loadfonts(device = "win")
viridis_colors <- viridis_pal()(3)
setwd("C:/Users/eb826/OneDrive - University of Exeter/PhD_EdouardMicrosporidia/Thesis/Chapter4_Leidyana gryllorum/Density experiment")
```

# Gametocysts analysis

## Import data

```{r }
gametocysts_data <- read_excel("Gametocysts_Feces_data.xlsx")
colnames(gametocysts_data)
```

## Convert factors to appropriate levels

```{r }
gametocysts_data[c("Treatment", "Infection", "Density", "R", "Dpi", "Box_Id", "Cricket_Age")] <- lapply(gametocysts_data[c("Treatment", "Infection", "Density", "R", "Dpi", "Box_Id", "Cricket_Age")], factor)
```

## Reorganise levels of Treatment variable and convert variables to factors

```{r }
gametocysts_data$Treatment <- factor(gametocysts_data$Treatment,levels = c("N_10","Y_10","N_30","Y_30","N_70","Y_70"))
```

## Data summary

```{r }
str(gametocysts_data)
summary(gametocysts_data)

# Calculate summary statistics for each treatment group
summary_stats <- aggregate(Gametocysts ~ Treatment, data = gametocysts_data, 
                           FUN = function(x) c(mean = mean(x), 
                                               median = median(x), 
                                               sd = sd(x), 
                                               min = min(x), 
                                               max = max(x)))
print(summary_stats)
```

## Create boxplots for each treatment group

```{r}
Gametocysts_boxplot <- ggplot(gametocysts_data, aes(x = Treatment, y = Gametocysts, color = Treatment)) +
  geom_boxplot() +
  geom_point() +
  scale_color_viridis_d(labels = c("N_10","Y_10","N_30","Y_30","N_70","Y_70")) +
  theme_bw(base_family = "Times New Roman") +
  theme(
    axis.text = element_text(size = 12), 
    axis.title = element_text(size = 12, vjust = 0.5),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.direction = "vertical",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5)) +
  labs(title = "Gametocysts per Treatment",
       color = "Infection_Density")
Gametocysts_boxplot
#ggsave("Gametocysts_boxplot.png", Gametocysts_boxplot, dpi = 500, width = 6, height = 4, units = "in")
```

## Create gametocyst curves over time

```{r}
# Calculate mean Gametocysts per dpi obtained by treatment
mean_gametocysts <- aggregate(Gametocysts ~ Dpi + Treatment, data = gametocysts_data, FUN = mean)

# Plot with mean Gametocysts per dpi and curves for each treatment
Gametocyst_curves_dpi <- ggplot(mean_gametocysts, aes(x = Dpi, y = Gametocysts, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, shape = 15) +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000")) +
  theme_bw(base_family = "Times New Roman") +
  theme(
    axis.text = element_text(size = 12), 
    axis.title = element_text(size = 12, vjust = 0.5),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.direction = "vertical",
    legend.position = "right",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    ) +
  labs(title = "Mean Gametocysts per Dpi and Treatment",
       x = "Dpi",
       y = "Mean Gametocysts",
       color = "Infection_Density")
Gametocyst_curves_dpi

# Calculate mean Gametocysts per cricket age obtained by treatment
mean_gametocysts_age <- aggregate(Gametocysts ~ Cricket_Age + Treatment, data = gametocysts_data, FUN = mean)

# Define the new rows to add
new_rows <- data.frame(
  Cricket_Age = factor(rep(15, 6)), # Assign Cricket Age as 15 and convert to factor
  Treatment = c("N_10", "Y_10", "N_30", "Y_30", "N_70", "Y_70"), # Adjust treatment names accordingly
  Gametocysts = rep(0, 6) # Assign Gametocysts as 0
)

# Combine factor levels from the existing and new rows
updated_mean_gametocysts <- rbind(mean_gametocysts_age, new_rows)
# Reorder the levels of Cricket_Age
updated_mean_gametocysts$Cricket_Age <- factor(updated_mean_gametocysts$Cricket_Age, levels = c("15", "19", "22", "26", "29", "33", "36"))

# Add text for infection rounds
y <- max(mean_gametocysts_age$Gametocysts) * 0.95
textaes <- data.frame(y = c(y, y, y),
                      x = c(1, 3, 5),
                      lab = c("Infection 1", "Infection 2", "Infection 3"))

# Plot with mean Gametocysts per cricket age and curves for each treatment
Gametocyst_curves_age <- ggplot(updated_mean_gametocysts, aes(x = Cricket_Age, y = Gametocysts, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 2, shape = 15) +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000")) +
  theme_bw(base_family = "Times New Roman") +
  theme(
    axis.text = element_text(size = 12), 
    axis.title = element_text(size = 12, vjust = 0.5),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.direction = "horizontal",
    legend.position = "top",
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    ) +
  labs(title = "Mean Gametocysts per Cricket Age and Treatment",
       x = "Cricket Age (days)",
       y = "Mean Gametocysts",
       color = "Infection_Density") +
  geom_segment(aes(x = 1, y = 0, yend = max(mean_gametocysts_age$Gametocysts) * 0.9), 
               colour = "red", linetype = "dashed", linewidth = 0.5) + 
  geom_segment(aes(x = 3, y = 0, yend = max(mean_gametocysts_age$Gametocysts) * 0.9), 
               colour = "red", linetype = "dashed", linewidth = 0.5) +
  geom_segment(aes(x = 5, y = 0, yend = max(mean_gametocysts_age$Gametocysts) * 0.9), 
               colour = "red", linetype = "dashed", linewidth = 0.5) + # Add infection rounds at age 15, 22 and 29 
   geom_text(mapping = aes(y = y, x = x, label = lab), colour = "red", data = textaes, inherit.aes = FALSE)

Gametocyst_curves_age
```

## Correlation between gametocysts and other parameters

### Plot

```{r }
correlation_plot1 <- ggplot(gametocysts_data, aes(x = Feces, y = Gametocysts)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Correlation between gametocysts and faeces",
       x = "Faeces per box (mg)",
       y = "Gametocysts count per box") +
  theme_bw(base_family = "Times New Roman") + 
 theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),  
        axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12),  
        legend.position = "none")
correlation_plot1

correlation_plot2 <- ggplot(gametocysts_data, aes(x = Feces, y = Gametocysts)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Correlation between gametocysts and faeces",
       x = "Faeces per box (mg)",
       y = "Gametocysts count per box") +
  theme_bw(base_family = "Times New Roman") + 
 theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),  
        axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12),  
        legend.position = "none") +
  facet_wrap(~ Dpi, scales = "free")
correlation_plot2
```

### Compute correlation coefficient

```{r }
correlation <- cor(gametocysts_data$Feces, gametocysts_data$Gametocysts)
# Conduct Pearson's correlation test
cor_test <- cor.test(gametocysts_data$Feces, gametocysts_data$Gametocysts)
# Print correlation coefficient
print(paste("Correlation coefficient:", correlation))
# Print correlation test results
print(cor_test)

# Compute correlation for each Dpi level
correlation_by_dpi <- gametocysts_data %>%
  group_by(Dpi) %>%
  summarise(correlation = cor(Gametocysts, Feces))
print(correlation_by_dpi)

# Correlation tests for each Dpi level
correlation_test_by_dpi <- gametocysts_data %>%
  group_by(Dpi) %>%
  group_map(~ cor.test(.x$Gametocysts, .x$Feces))

library(purrr)
library(broom)
correlation_test_summary <- map(correlation_test_by_dpi, tidy)
correlation_test_summary <- bind_rows(correlation_test_summary)
print(correlation_test_summary)
```

## General model

```{r }
# Look at the variable distrib
hist(gametocysts_data$Gametocysts)

# ziformula = ~Treatment Assuming the covariate Treatment is included in the zero-inflation part
 
model_gameto_age_nb <- glmmTMB(Gametocysts ~  Infection * Density + Cricket_Age + Feces + (1|Box_Id), data = gametocysts_data, family = nbinom2(link = "log"))

model_gameto_dpi_nb <- glmmTMB(Gametocysts ~  Infection * Density + Dpi + Feces + (1|Box_Id), data = gametocysts_data, family = nbinom2(link = "log"))

anova(model_gameto_age_nb, model_gameto_dpi_nb)
# model_gameto_age_nb has a better fit
```

### Check model

```{r }
library(MuMIn)
options(na.action = "na.fail") # prevent fitting models to different datasets, needed for dredge()
dredge(model_gameto_age_nb)
# The model with all the interactions has the better fit

# Selcted final for analysis
model_gameto_age_final <- glmmTMB(Gametocysts ~  Infection * Density + Cricket_Age + Feces + (1|Box_Id), data = gametocysts_data, family = nbinom2(link = "log"))
library(performance)
check_overdispersion(model_gameto_age_final)
# Overdispersion test: dispersion ratio = 0.469 (p = 0.152)

summary(model_gameto_age_final)
Anova(model_gameto_age_final, type="II")

pairwise_comp_i <- emmeans(model_gameto_age_final, pairwise ~ Infection, type="response")
pairwise_comp_i
cld(pairwise_comp_i,adjust="fdr",Letters=letters)

pairwise_comp_d <- emmeans(model_gameto_age_final, pairwise ~ Density, type="response")
pairwise_comp_d
cld(pairwise_comp_d,adjust="fdr",Letters=letters)

pairwise_comp_age <- emmeans(model_gameto_age_final, pairwise ~ Cricket_Age, type="response")
pairwise_comp_age
cld(pairwise_comp_age,adjust="fdr", Letters=letters)

plot_age_emmean <- plot(pairwise_comp_age, horizontal=FALSE, comparisons=TRUE) +
        theme_bw(base_family = "Times New Roman") + 
        theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),  
        axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, size = 12)) +
  labs(title = "Emmeans of gametocysts",
       x = "Cricket age (days)",
       y = "Gametocyst emmeans")
plot_age_emmean
```

# Survival analysis

## Import data

```{r }
survival_data <- read_excel("SurvivalTable.xlsx")
colnames(survival_data)
```
## Convert factors to appropriate levels

```{r }
survival_data[c("Treatment", "Infection", "Density", "R")] <- lapply(survival_data[c("Treatment", "Infection", "Density", "R")], factor)
```

## Reorganise levels of Treatment variable and convert variables to factors

```{r }
survival_data$Treatment <- factor(survival_data$Treatment,levels = c("N_10","Y_10","N_30","Y_30","N_70","Y_70"))
```

## Data summary

```{r }
str(survival_data)
summary(survival_data)
```

## Check mortality (sanity check)

```{r }
# Calculate total number of crickets and number of deaths by treatment
total_Adomesticus <- survival_data %>% 
  group_by(Density) %>% 
  summarise(total = n())

mortality_Adomesticus <- survival_data %>% 
  group_by(Density, Censor) %>% 
  summarise(mortality = sum(Censor == 1)) %>%
  filter(Censor == 1) # Keep only dead crickets

# Merge total and mortality data frames
Adomesticus_check <- merge(total_Adomesticus, mortality_Adomesticus, by = "Density", all = TRUE)

# Calculate percentage of mortality
Adomesticus_check$percent <- (Adomesticus_check$mortality / Adomesticus_check$total) * 100
print(Adomesticus_check)

# Create a data frame with the required columns
contingency_data <- data.frame(Density = Adomesticus_check$Density,
                               Alive = Adomesticus_check$total - Adomesticus_check$mortality,
                               Dead = Adomesticus_check$mortality)

# Set Density as row names
row.names(contingency_data) <- contingency_data$Density
contingency_data$Density <- NULL

# Reorder rows according to desired order
desired_order <- c("10","30","70")
contingency_data <- contingency_data[match(desired_order, rownames(contingency_data)), ]
print(contingency_data)

# Perform chi-square test
chi_square_test <- chisq.test(contingency_data)
print(chi_square_test)

# Perform pairwise comparisons using fdr correction
contingency_matrix <- as.matrix(contingency_data)
pairwise_chisq <- pairwise.prop.test(x = contingency_matrix, p.adjust.method = "fdr")
print(pairwise_chisq)
```

## Surv curves by treatments

```{r }
# survival_data$Death_time <- survival_data$Death_time + 15
# Calculate the labels for the new Death_time
#labs_new <- paste("New Death Time:", survival_data$Death_time)

surv <- survfit(Surv(Death_time, Censor) ~ Density, data = survival_data)
surv_plot <- ggsurvplot(
  fit = surv,
  ylim = c(0.7, 1),
  surv.scale = "percent",
  title = "Cricket survival per density treatment",
  xlab ="Cricket age (days)",
  legend.title = "Density",
  legend.labs = c("10","30","70"),
  palette = viridis_colors,  # Use viridis color palette
  #palette = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000"),
  #linetype = c(1, 2, 3),
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10, face = "bold"),
      legend.direction = "horizontal",
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot
#ggsave("surv_plot.png", surv_plot$plot, dpi = 300, width = 6, height = 4, units = "in")

#annotate(geom = "text", x = survival_data$Death_time, label = labs_new, size=4)

# Get p-values with coordinates
#pvalue_letters <- surv_pvalue(surv, method = "survdiff", get_coord = TRUE)
```

## LT50 determined with median values

```{r }
summary(surv)$table
```

## Determine if we have a constant risk or not

```{r }
plotsurvivors(survival_data$Death_time, survival_data$Censor)
# The instantaneous risk is considered constant if the points on the survival curve are roughly aligned on a straight line.
## Here the instant risk is constant, then we opt for a survival regression model
```

## Surv reg model

### Find the best distribution for our survival data
```{r}
# Create vector of parametric distributions to test
distList <- c(distList <- c("extreme", "logistic", "gaussian", "weibull", "exponential", "rayleigh", "lognormal", "loglogistic", "t"))

# Function fits each distribution and extracts AIC, BIC, log-likelihood values
## The distribution is estimated on the event data without any covariate (~ 1)
fit_dist <- function(dist) {
  tmp <- survreg(Surv(Death_time, Censor) ~ 1, data = survival_data, dist = dist)
  c(AIC(tmp), BIC(tmp), as.numeric(logLik(tmp)))
}

# Apply above function to each distribution in the distList
results_list <- lapply(distList, fit_dist)

# Convert the above list of results to a data frame
results_df <- data.frame(t(matrix(unlist(results_list), nrow = 3)))
colnames(results_df) <- c("aic", "bic", "logLik")
rownames(results_df) <- distList

# Find the distribution with the lowest AIC, BIC, and logLik values
bestFitAIC <- rownames(results_df)[which.min(results_df$aic)]
bestFitBIC <- rownames(results_df)[which.min(results_df$bic)]
bestFitLogLik <- rownames(results_df)[which.max(results_df$logLik)]

# Print the results data frame and best fitting distributions
results_df
cat("\nBest fitting distribution using AIC:", bestFitAIC, "\n")
cat("\nBest fitting distribution using BIC:", bestFitBIC, "\n")
cat("\nBest fitting distribution using Log-Likelihood:", bestFitLogLik, "\n")
## The best distribution here is lognormal 

# We can check graphically how the distribution is fitting to the survival data
require(flexsurv)
s <- with(survival_data,Surv(Death_time,Censor))
sLnorm  <- flexsurvreg(s ~ 1,dist="lnorm",data=survival_data)
sExp  <- flexsurvreg(s ~ 1,dist="exponential",data=survival_data) 
plot(sLnorm, ylim = c(0.7, 1)) +
  lines(sExp, col = "blue")
```

```{r }
model_surv <- survreg(Surv(Death_time, Censor) ~ Density, data = survival_data, dist="lognormal") # distribution determined earlier
RVAideMemoire::plotresid(model_surv)

summary(model_surv)
Anova(model_surv, type="II")
```

## Pairwise differences between different treatments

```{r }
pairwise_comp_sd <- emmeans(model_surv, pairwise ~ Density, type="response")
pairwise_comp_sd
cld(pairwise_comp_sd,adjust="fdr",Letters=letters)
```

# Emergence Analysis

## Import data

```{r }
emergence_data <- read_excel("EmergencelTable.xlsx")
colnames(emergence_data)
```

### Convert factors to appropriate levels

```{r }
emergence_data[c("Treatment","Infection","Density","R","Sex")] <- lapply(emergence_data[c("Treatment","Infection","Density","R","Sex")],factor)
```

### Reorganise levels of Treatment variable and convert variables to factors

```{r }
emergence_data$Treatment <- factor(emergence_data$Treatment,levels = c("N_10","Y_10","N_30","Y_30","N_70","Y_70"))
```

### Data summary

```{r }
str(emergence_data)

# Filter the data to include only rows where Censor equals 1
filtered_data <- filter(emergence_data, Censor == 1)

# Group by Treatment, Infection, and Density, and summarize the counts of Sex
emmergence_summary_table <- filtered_data %>%
  group_by(Treatment, Infection, Density, R) %>%
  summarise(Count_of_Sex = n()) %>%
  arrange(Treatment, Infection, Density, R)

# Print the summarized table
print(emmergence_summary_table)
```

## Calculate proportions and perform chi-square tests

```{r echo=TRUE, message=FALSE}
emergence_table <- emergence_data %>%
  group_by(Density) %>%
  summarise(y1 = sum(Censor == 1), #Number of emerged crickets
            y2 = sum(Censor == 0)   #Number of non-emerged crickets
  ) %>%
  mutate(y1_prop = y1 / (y1 + y2), #Proportion of emerged crickets
         y2_prop = y2 / (y1 + y2))   #Proportion of non-emerged crickets

print(emergence_table)

# Create the contingency table
contingency_table <- as.matrix(emergence_table[, c("y1", "y2")])
rownames(contingency_table) <- paste(emergence_table$Density)
colnames(contingency_table) <- c("Emerged", "Non-emerged")

# Perform chi-square test
chisq_result <- chisq.test(contingency_table)
print(chisq_result)

# Perform pairwise comparisons using fdr correction
pairwise_chisq <- pairwise.prop.test(x = contingency_table, p.adjust.method = "fdr")
print(pairwise_chisq)
```
## LT50 table

```{r }
surv_emergence <- survfit(Surv(Emergence_time,Censor) ~ Sex,data = emergence_data)
summary(surv_emergence)$table
```

## Surv curve emergence

```{r }
surv_emergence <- survfit(Surv(Emergence_time,Censor) ~ Density,data = emergence_data)
surv_plot_emergence <- ggsurvplot(
  fit = surv_emergence,
  surv.scale = "percent",
  title = "Cricket emergence across density treatment",
  ylab = "Cricket emergence",
  xlab = "Cricket age (days)",
  legend.title = "Density",
  legend.labs = c("10","30","70"),
  palette = viridis_colors,
  ylim=c(0, 1),
  break.y.by = 0.2,
  xlim = c(36,53),
  break.x.by = 2,
  fun="event",
  #surv.median.line = "h",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10, face = "bold"),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))

# Add line at 50% emergence
p1 <- surv_plot_emergence$plot +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gold")
p1
#ggsave("emergence_plot.png", p1, dpi = 500, width = 4.5, height = 4, units = "in")
```

## Surv curve male

```{r }
# Subset for males
df_males <- subset(emergence_data, Sex == "Male" | Sex == "NA")


surv_male <- survfit(Surv(Emergence_time,Censor) ~ Density,data = df_males)
surv_plot_male <- ggsurvplot(
  fit = surv_male,
  surv.scale = "percent",
  title = "Cricket male emergence",
  ylab = "Cricket emergence",
  palette = viridis_colors,
  legend = "none",
  xlim = c(38,53),
  break.x.by = 5,
  fun="event",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot_male
#ggsave("emergence_male_plot.png", surv_plot_male$plot, dpi = 500, width = 2.6, height = 3.5, units = "in")
```

## Surv curve female

```{r }
# Subset for females
df_females <- subset(emergence_data, Sex == "Female" | Sex == "NA")

surv_female <- survfit(Surv(Emergence_time,Censor) ~ Density,data = df_females)
surv_plot_female <- ggsurvplot(
  fit = surv_female,
  surv.scale = "percent",
  title = "Cricket female emergence",
  ylab = "Cricket emergence",
  palette = viridis_colors,
  legend = "none",
  xlim = c(36,53),
  break.x.by = 5,
  #ylim = c(0,0.4),
  fun="event",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot_female
#ggsave("emergence_female_plot.png", surv_plot_female$plot, dpi = 500, width = 2.6, height = 3.5, units = "in")
```

## Arrange the surv plots together

```{r }
splots <- list(surv_plot_female, surv_plot_male)
arrange <- arrange_ggsurvplots(splots, print = TRUE, ncol = 2, nrow = 1)
arrange
```

## Generalized linear model – binomial family

```{r }
# Filter out NA values from the initial data set
emergence_data_filtered <- emergence_data %>% filter(Sex != "NA")

glm_model <- glm(Emergence_time ~ Density + Sex, data = (emergence_data),family = Gamma(link = "inverse"))
```

### Check model and Anova

```{r }
RVAideMemoire::plotresid(glm_model)

summary(glm_model)
Anova(glm_model,type="II")
# Only Sex is significant, with Infection and Infection:Density almost significant (0.09)
```

# Weight Analysis

## Import data

```{r }
weight_data <- read_excel("Weight_data.xlsx")
colnames(weight_data)
```

## Reorganise levels of Treatment variable and convert variables to factors

```{r }
## Reorganise levels of Treatment variable and convert variables to factors
weight_data <- weight_data %>%
  mutate(
    Treatment = factor(Treatment, levels = c("N_10","Y_10","N_30","Y_30","N_70","Y_70")),
    Sex = factor(Sex),
    Infection = factor(Infection),
    Density = factor(Density),
    Rep = factor(Rep),
    Date_emerge = factor(Date_emerge)
    #Date_emerge = as.Date(Date_emerge, format="%Y/%m/%d")  # Convert Date_emerge to date format
  )
str(weight_data)
```

## Calculate mean and standard deviation for weight variable by treatment

```{r echo=TRUE}
summary_data_w <- aggregate(Weight ~ Treatment * Sex, weight_data, FUN = function(x) c(mean = mean(x), sd = sd(x)))
print(summary_data_w)
```

## Box plot of Weight by Treatment

```{r }
boxplot_weight <- ggplot(weight_data, aes(x = Density, y = Weight, color = Density),
  legend.title = "Density",
  legend.labs = c("10","30","70")) +
  geom_boxplot(aes(middle = mean(Weight))) +
  geom_point() +
  scale_color_viridis_d() +
  labs(title = "Maturation weight by density treatment",
       x = "Treatment",
       y = "Weight (mg)") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 50, vjust = 0.6),
        legend.title = element_text(size = 10, face = "bold"),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
```

## Box Plot of weight Facet by Sex

```{r }
boxplot_faceted <- boxplot_weight + facet_wrap(~ Sex) +
  theme(strip.text.x = element_text(size=12)) +
  labs(colour = "Density")
boxplot_faceted
#ggsave("boxplot_weight.png", boxplot_faceted, dpi = 500, width = 7, height = 5, units = "in")
```

## General model

```{r }
# Look at variable distrib
hist(weight_data$Weight) # It seems like a normal distrib

model_weight <- lm(Weight ~ Density + Sex, data = weight_data)
model_weight_d <- lm(Weight ~ Density + Sex + Date_emerge, data = weight_data)
anova(model_weight, model_weight_d)
# A lower Residual Sum of Squares (RSS) indicates a better fit. Model 2 including the emergence date  has a significant lower RSS
```

### Check residuals and anova

```{r }
RVAideMemoire::plotresid(model_weight_d)

summary(model_weight_d)
Anova(model_weight_d, type="II")
# Density, Sex and Date_emerge are significant terms
```

### Pairwise comparisons

```{r }
pairwise_comp_w <- emmeans(model_weight_d, pairwise ~ Density, type="response")
pairwise_comp_w
cld(pairwise_comp_w, adjust="fdr", Letters=letters)

pairwise_comp_ws <- emmeans(model_weight_d, pairwise ~ Density|Sex, type="response")
pairwise_comp_ws
cld(pairwise_comp_ws, adjust="fdr", Letters=letters)

pairwise_comp_w1 <- emmeans(model_weight_d, pairwise ~ Sex, type="response")
pairwise_comp_w1
cld(pairwise_comp_w1, adjust="fdr", Letters=letters)

#pairwise_comp_w2 <- emmeans(model_weight_d, pairwise ~ Sex*Date_emerge, type="response")
#pairwise_comp_w2
#cld_w2 <- cld(pairwise_comp_w2, adjust="fdr", Letters=letters)

pairwise_comp_w3 <- emmeans(model_weight_d, pairwise ~ Date_emerge, type="response")
pairwise_comp_w3
cld_w3 <- cld(pairwise_comp_w3, adjust="fdr", Letters=letters)

# Plot the emmeans object
## Remove spaces in .group 
cld_w3$.group=gsub(" ", "", cld_w3$.group)

# Apply the nudge amount in geom_text
plot_cld_w3 <- ggplot(cld_w3, aes(x = Date_emerge, y = emmean, label = .group)) +
  geom_point(shape = 15, size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax =  upper.CL), width = 0.2, size = 0.7) +
  geom_text(aes(y = upper.CL),  # Set y = upper.CL
            nudge_y = 15,  # Adjust this value as needed
            size = 3,
            color='black') +
  theme_bw(base_family = "Times New Roman") + 
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),  
        axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12)) +
  labs(title = "Weight emmeans across cricket age",
       x = "Cricket age (days) ",
       y = "Emmeans weight (mg)")
plot_cld_w3
#ggsave("Weight_emmean_age.png", plot_cld_w3, dpi = 500, width = 7, height = 5, units = "in")
```

# Feed and feces analysis

## Import data

```{r }
feed_feces_data <- read_excel("Feed_Feces_data.xlsx")
colnames(feed_feces_data)
summary(feed_feces_data)
```

## Reorganise levels of Treatment variable and convert variables to factors

```{r }
feed_feces_data[c("Treatment", "R", "Box_Id", "Infection", "Density", "Dpi", "Cricket_Age")] <- lapply(feed_feces_data[c("Treatment", "R", "Box_Id", "Infection", "Density", "Dpi", "Cricket_Age")], factor)
feed_feces_data$Treatment <- factor(feed_feces_data$Treatment,
                                                 levels = c("N_10","Y_10","N_30","Y_30","N_70","Y_70"))
str(feed_feces_data)
```

## Calculate mean and standard deviation for each response variable by treatment

```{r }
summary_data_f <- aggregate(cbind(Feces_ind_day, Feed_ind_day) ~ Treatment + Dpi, feed_feces_data, FUN = function(x) c(mean = mean(x), sd = sd(x)))
print(summary_data_f)
```

## Boxplot of Feed Consumed by Treatment

```{r }
boxplot_feed <- ggplot(feed_feces_data, aes(x = Treatment, y = Feed_ind_day, color = Treatment),
  legend.title = "Infection_Density",
  legend.labs = c("N_10","Y_10","N_30","Y_30","N_70","Y_70")) +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Feed consumed by treatment",
       x = "Treatment",
       y = "Feed consumed per individual and per day (mg)") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
boxplot_feed
```

## Boxplot of Feed Consumed by Treatment and Dpi

```{r }
# Create boxplot with filtered data
boxplot_feed <- ggplot(feed_feces_data, aes(x = Treatment, y = Feed_ind_day, color = Treatment)) +
  geom_boxplot(size = 0.8) +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Feed consumed by treatment and Dpi",
       x = "Treatment",
       y = "Feed consumed per individual and per day (mg)",
       colour = "Infection_Density") +
  facet_wrap(~Dpi) +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 50, vjust = 0.6), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        legend.position = c(.85, .2))
boxplot_feed
```

## Boxplot of Feces Produced by Treatment

```{r }
boxplot_feces <- ggplot(feed_feces_data, aes(x = Treatment, y = Feces_ind_day, color = Treatment),
  legend.title = "Infection_Density",
  legend.labs = c("N_10","Y_10","N_30","Y_30","N_70","Y_70")) +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Faeces produced by treatment",
       x = "Treatment",
       y = "Faeces per individual and per day (mg)") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
boxplot_feces
```

## Boxplot of Feces by Treatment and Dpi

```{r }
boxplot_feces <- ggplot(feed_feces_data, aes(x = Treatment, y = Feces_ind_day, color = Treatment)) +
  geom_boxplot(aes(middle = mean(Feces_ind_day)), size = 0.8) +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Faeces produced by treatment and Dpi",
       x = "Treatment",
       y = "Faeces produced per individual and per day (mg)",
       colour = "Infection_Density") +
  facet_wrap(~Dpi) +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 50, vjust = 0.6), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        legend.position = c(.85, .2))
boxplot_feces
```

## Correlation between consumed feed and Faeces

```{r }
correlation_plot <- ggplot(feed_feces_data, aes(x = Feces_ind_day, y = Feed_ind_day)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Correlation between consumed feed and faeces",
       x = "Faeces per individual and per day (mg)",
       y = "Feed consumed per individual and per day (mg)") +
  theme_bw(base_family = "Times New Roman") + 
 theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),  
        axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12),  
        legend.position = "none")

correlation_plot
#ggsave("correlation_plot.png", correlation_plot, dpi = 500, width = 7, height = 5, units = "in")
```

### Compute correlation coefficient

```{r }
correlation <- cor(feed_feces_data$Feed_ind_day, feed_feces_data$Feces_ind_day)
```

### Conduct Pearson's correlation test

```{r }
cor_test <- cor.test(feed_feces_data$Feed_ind_day, feed_feces_data$Feces_ind_day)
```

### Print correlation coefficient

```{r }
print(paste("Correlation coefficient:", correlation))
```

### Print correlation test results

```{r }
print(cor_test)
# There is a strong positive correlation between the amount of feed consumed and the amount of feces produced
```

## Linear Model (lm) for biomass and Quantitative continuous variables

### Feces model

```{r }
hist(feed_feces_data$Feces_ind_day)

model_feces <- lmer(Feces_ind_day ~  Density + (1|Box_Id), data = feed_feces_data) #The box id is a random effect (different box measured across different time points)
model_feces_age <- lmer(Feces_ind_day ~  Density + Cricket_Age + (1|Box_Id), data = feed_feces_data)


anova(model_feces, model_feces_age)
# The model including Cricket age has significant lower AIC and BIC and a higher logLik
```

#### Check residuals and compare the models and test significance of each term of the model with Anova

```{r }
RVAideMemoire::plotresid(model_feces_age)

summary(model_feces_age)
Anova(model_feces_age, type="II")
```

#### Pairwise comparisons

```{r }
pairwise_comp_faeces <- emmeans(model_feces_age, pairwise ~ Density, type="response")
pairwise_comp_faeces
cld(pairwise_comp_faeces, adjust="fdr", Letters=letters)

pairwise_comp_faeces2 <- emmeans(model_feces_age, pairwise ~ Cricket_Age, type="response")
pairwise_comp_faeces2
cld(pairwise_comp_faeces2, adjust="fdr", Letters=letters)
```

# Create a table that compiles all the above statistical outputs from the Anova objects

* Function to extract results

```{r}
get.univariate.results <- function(model) {
  anova_results <- car::Anova(model, type = "II") %>% broom::tidy()
  
  if ("lme" %in% class(model) || "lmerMod" %in% class(model) || "glmerMod" %in% class(model)) {
    random_effects <- paste(names(VarCorr(model)), collapse = ", ")
    variances <- paste(format(VarCorr(model), digits = 2), collapse = ", ")
  } else {
    random_effects <- NA
    variances <- NA
  }
  
  cbind(
    `Response (y)` = deparse(summary(model)$call$formula[[2]]),
    `Model` = as.character(summary(model)$call[[1]]),
    Formula = paste ("y ~", deparse(summary(model)$call$formula[[3]])),
    anova_results %>%
      dplyr::mutate(
        statistic = round(statistic, digits = 2),
        signif = gtools::stars.pval(p.value),
        p.value = round(p.value, digits = 3),
        p.value = format.pval(p.value, digits = 3)
      ),
    random_effects = random_effects,
    variances = variances
  ) %>% as.data.frame
}
```

* Create a list that contains all models

```{r}
sum.model <- list(model_gameto_age_final, glm_model, model_weight_d, model_feces_age)
```

* Apply the function to the list

```{r message=FALSE, warning=FALSE}
univariate.output <- purrr::map(.x = sum.model, .f = get.univariate.results) %>%
  dplyr::bind_rows()
univariate.output <- dplyr::rename(univariate.output, c(Factor = term, `Test value` = statistic, `P value`= p.value))

knitr::kable(univariate.output)
write.table(univariate.output, file = "models_stat_densityexp.txt", sep = ";", dec = ".")
```